### 2016-03-03 docs/CDC 2015 platooning/arXiv/solution.tex
no message

\noindent where $x\in\mathbb{R}^n$ is the system state, $u_1\in \mathcal{U}_1$ is the control of Player 1, and $u_2\in\mathcal{U}_2$ is the control of Player 2. We assume $f:\mathbb{R}^n\times \mathcal{U}_1 \times \mathcal{U}_2 \rightarrow \mathbb{R}^n$ is uniformly continuous, bounded, and Lipschitz continuous in $x$ for fixed $u_1,u_2$, and the control functions $u_1(\cdot)\in\mathbb{U}_1,u_2(\cdot)\in\mathbb{U}_2$ are drawn from the set of measurable functions. Player 2 is allowed to use nonanticipative strategies \cite{Evans84,Varaiya67} $\gamma$, defined by

@@

\begin{equation}
\begin{aligned}
\mathcal{V}(t) &:= \{x\in\mathbb{R}^n \mid \exists \gamma\in\Gamma \text{ such that } \forall u_1(\cdot)\in\mathbb{U}_1, \\
&\exists s \in [t,0], \xi_f(s; t, x, u_1(\cdot), \gamma[u_1](\cdot)) \in \mathcal{L} \}
\end{aligned}
\end{equation}
where $\xi_f$ is the trajectory of the system satisfying initial conditions $\xi_f(t; x, t, u_1(\cdot), u_2(\cdot))=x$ and the following differential equation almost everywhere on $[-t, 0]$:
\begin{equation}
\begin{aligned}
\frac{d}{ds}&\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)) \\
&= f(\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)), u_1(s), u_2(s))
\end{aligned}
\end{equation}

For this paper, we use the HJ formulation in \cite{Mitchell05}, which has shown that the backwards reachable set $\mathcal{V}(t)$ can be obtained as the zero sublevel set of the viscosity solution \cite{Crandall84} $V(t,x)$ of the following terminal value Hamilton-Jacobi-Isaacs (HJI) PDE:

\begin{equation} \label{eq:HJIPDE}
\begin{aligned}
D_t &V +\min \{0, \max_{u_1\in\mathcal{U}_1} \min_{u_2\in\mathcal{U}_2} D_x V \cdot f(x,u_1,u_2) \} = 0, \\
&V(0,x) = l(x)
\end{aligned}
\end{equation}

@@

\section{Liveness Controllers \label{sec:liveness}}
\subsection{Merging onto a Highway \label{subsec:highway_merge}}
We model the merging of a vehicle onto an air highway as a path planning problem, where we specify a target position and velocity along the highway. Thus, a vehicle would aim to drive the system \eqref{eq:dyn} to a state $\bar{x}_H=(\bar{p}_x, \bar{v}_x, \bar{p}_y, \bar{v}_y)$, or a small range of states defined by the set

@@

In this reachability problem, $\mathcal{L}_H$ is the target set, represented by the zero sublevel set of the function $l_H(x)$, which specifies the terminal condition of the HJB PDE to be solved. The solution we obtain, $V_H(t,x)$, is the implicit surface function representing the reachable set $\mathcal V_H(t)$; $V_H(-T,x)\le 0$, then, specifies the reachable set $\mathcal{V}_H(T)$, the set of states from which the system can be driven to the target $\mathcal{L}_H$ within a duration of $T$. This gives the algorithm for merging onto the highway:

\begin{enumerate}
\item Move towards $\bar{x}_H$ in a straight line until $V_H(-T,x)\le 0$. This simple controller is chosen heuristically.
\item Apply the optimal control extracted from $V_H(-T,x)$ according to \eqref{eq:HJB_ctrl_syn} until $\mathcal{L}_H$ is reached.
\end{enumerate}

\subsection{Merging into a Platoon \label{subsec:platoon_merge}}
We again pose the merging of a vehicle into a platoon on an air highway as a reachability problem. Here, we would like quadrotor $Q_i$ to merge onto the highway and follow another vehicle $Q_j$ in a platoon. Thus, we would like to drive the system given by \eqref{eq:rel_dyn_aug} to a specific $\bar{x}_P = (\bar{p}_{x,r}, \bar{v}_{x,r}, \bar{p}_{y,r}, \bar{v}_{y,r})$, or a small range of relative states defined by the set

@@

The target set $\mathcal{L}_P$ is represented by the implicit surface function $l_P(x)$, which specifies the terminal condition of the HJI PDE \eqref{eq:HJIPDE}. The zero sublevel set of the solution to \eqref{eq:HJIPDE}, $V_P(-T,x)$, gives us the set of relative states from which $Q_i$ can reach the target and join the platoon following $Q_j$ within a duration of $T$. We assume that $Q_j$ moves along the highway at constant speed, so that $u_j(t)$ = 0. The following is a suitable algorithm for a vehicle merging onto a highway and joining a platoon to follow $Q_j$:

\begin{enumerate}
\item Move towards $\bar{x}_P$ in a straight line until $V_P(-T,x)\le 0$.
\item Apply the optimal control extracted from $V_P(-T,x)$ according to \eqref{eq:HJI_ctrl_syn} until $\mathcal{L}_P$ is reached.
\end{enumerate}

\subsection{Other Quadrotor Maneuvers}
For the simpler maneuvers of traveling along a highway and following a platoon, we resort to simpler controllers described below.

\subsubsection{Traveling along a highway} \label{sec:travel_hwy}
We use a model-predictive controller (MPC) for traveling along a highway at a pre-specified speed. Here, a leader quadrotor tracks a constant-altitude path, defined as a curve $\bar{p}(s)$ parametrized by $s\in[0,1]$ in $p=(p_x, p_y)$ space (position space), while maintaining a velocity $\bar{v}(s)$ that corresponds to constant speed in the direction of the highway.

\subsubsection{Following a Platoon} \label{sec:follow_platoon}
Follower vehicles use a feedback control law tracking a nominal position and velocity in the platoon, with an additional feed-forward term given by the leader's acceleration input.

The $i$-th member of the platoon, $Q_{P_i}$, is expected to track a relative position in the platoon $r^i = (r_x^i,r_y^i)$ with respect to the leader's position $p_{P_1}$, and the leader's velocity $v_{P_1}$ at all times. The resulting control law has the form:
\begin{equation}\label{eq:follow}
u^i(t) = k_p \big[p_{P_1}(t) + r^i(t) - p^i(t) \big] + k_v\big[v_{P_1}(t) - v^i(t)\big] + u_{P_1}(t)
\end{equation}
for some $k_p,k_v>0$. A simple rule for determining $r^i(t)$ in a single-file platoon is given for $Q_{P_i}$ as:
\begin{equation}\label{eq:nominal_pos}
r^i(t) = - (i-1) b \frac{v_{P_1}}{\|v_{P_1}\|_2}
\end{equation}
where $b$ is the spacing between vehicles along the platoon. and $\frac{v_{P_1}}{\|v_{P_1}\|_2}$ is the platoon leader's direction of travel.

\section{Safety Controllers \label{sec:safety}}
\subsection{Wrapping Reachability Around Existing Controllers}
A quadrotor can use a liveness controller when it is not in any danger of collision with other quadrotors or obstacles. If the quadrotor could potentially be involved in a collision within the next short period of time, it must switch to a safety controller. In this section, we will demonstrate how HJ reachability can be used to both detect imminent danger and synthesize a controller that guarantees safety within a specified time horizon. For our safety analysis, we will use the model in \eqref{eq:rel_dyn_aug}.

We begin by defining the target set $\mathcal{L}_S$, which characterizes configurations in relative coordinates for which vehicles $Q_i,Q_j$ are considered to be in collision:

@@

\subsection{Platoon Safety Guarantees \label{subsec:safety_guarantees}}
Under normal operations in a single platoon, each follower quadrotor $Q_{P_i},i>1$ checks whether it is in the safety reachable set with respect to $Q_{P_{i-1}}$ and $Q_{P_{i+1}}$. So $\mathcal{Q}(i) = \{P_{i+1}, P_{i-1}\}$ for $i=P_2,\ldots,P_{N-1}$. Assuming there are no nearby quadrotors outside of the platoon, the platoon leader $Q_{P_1}$ checks safety against $Q_{P_2}$, and the platoon trailer $Q_{P_N}$ checks safety against $Q_{P_{N-1}}$. So $\mathcal{Q}(P_1)=\{P_2\}, \mathcal{Q}(P_N)=\{P_{N-1}\}$. No pair of quadrotors should be in an unsafe configuration if the liveness controllers are well-designed. Occasionally, a quadrotor $Q_k$ may behave unexpectedly due to faults, which may lead to an unsafe configuration.

With our choice of $\mathcal{Q}(i)$ and the assumption that the platoon is in a single-file formation, some quadrotor $Q_i$ would get into an unsafe configuration with $Q_k$, where $Q_k$ is likely to be the quadrotor in front or behind of $Q_i$. In this case, a ``safety breach" occurs. Our synthesis of the safety controller guarantees that between every pair of quadrotors $Q_i,Q_k$, as long as $V_S(-t,x_i- x_k)>0$, $\exists u_i$ to keep $Q_i$ from colliding with $Q_k$ for a desired time horizon $t$, despite the worst case (an adversarial) control from $Q_k$. Therefore, as long as the number of ``safety breaches" is at most one, $Q_i$ can simply use the optimal control $u_i$ to avoid collision with $Q_k$ for the time horizon of $t$. Since by assumption, vehicles in platoons are able to exit the current altitude range within a duration of $t_\text{internal}$, if we choose $t=t_\text{internal}$, the safety breach would always end before any collision can occur.

>>>>

\noindent where $x\in\mathbb{R}^n$ is the system state, $u_1\in \mathcal{U}_1$ is the control of Player 1, and $u_2\in\mathcal{U}_2$ is the control of Player 2. We assume $f:\mathbb{R}^n\times \mathcal{U}_1 \times \mathcal{U}_2 \rightarrow \mathbb{R}^n$ is uniformly continuous, bounded, and Lipschitz continuous in $x$ for fixed $u_1,u_2$, and the control functions $u_1(\cdot)\in\mathbb{U}_1,u_2(\cdot)\in\mathbb{U}_2$ are drawn from the set of measurable functions\footnote{
A function $f:X\to Y$ between two measurable spaces $(X,\Sigma_X)$ and $(Y,\Sigma_Y)$ is said to be measurable if the preimage of a measurable set in $Y$ is a measurable set in $X$, that is: $\forall V\in\Sigma_Y, f^{-1}(V)\in\Sigma_X$, with $\Sigma_X,\Sigma_Y$ $\sigma$-algebras on $X$,$Y$.}. Player 2 is allowed to use nonanticipative strategies \cite{Evans84,Varaiya67} $\gamma$, defined by

@@

\begin{equation}
\begin{aligned}
\mathcal{V}(t) &:= \{x\in\mathbb{R}^n \mid \exists \gamma\in\Gamma \text{ such that} \forall u_1(\cdot)\in\mathbb{U}_1, \\
&\exists s \in [t,0], \xi_f(s; t, x, u_1(\cdot), \gamma[u_1](\cdot)) \in \mathcal{L} \}
\end{aligned}
\end{equation}
where $\xi_f$ is the trajectory of the system satisfying initial conditions $\xi_f(t; x, t, u_1(\cdot), u_2(\cdot))=x$ and the following differential equation almost everywhere on $[-t, 0]$
\begin{equation}
\begin{aligned}
\frac{d}{ds}&\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)) \\
&= f(\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)), u_1(s), u_2(s))
\end{aligned}
\end{equation}

Many methods involving solving HJ PDEs \cite{Mitchell05} and HJ variational inequalities (VI) \cite{Bokanowski10,Barron89,Fisac15} have been developed for computing the reachable set. These HJ PDEs and HJ VIs can be solved using well-established numerical methods. For this paper, we use the formulation in \cite{Mitchell05}, which has shown that the backwards reachable set $\mathcal{V}(t)$ can be obtained as the zero sublevel set of the viscosity solution \cite{Crandall84} $V(t,x)$ of the following terminal value Hamilton-Jacobi-Isaacs (HJI) PDE:

\begin{equation} \label{eq:HJIPDE}
\begin{aligned}
D_t &V(t,x) + \\
&\min \{0, \max_{u_1\in\mathcal{U}_1} \min_{u_2\in\mathcal{U}_2} D_x V(t,x) \cdot f(x,u_1,u_2) \} = 0, \\
&V(0,x) = l(x)
\end{aligned}
\end{equation}

@@

\subsection{Relative Dynamics and Augmented Relative Dynamics}
Besides Equation \eqref{eq:dyn}, we will also consider the relative dynamics between two quadrotors $Q_i,Q_j$. These dynamics can be obtained by defining the relative variables

\begin{equation} \label{eq:rel_var}
\begin{aligned}
p_{x,r} &= p_{x,i} - p_{x,j}, \qquad p_{y,r} = p_{y,i} - p_{y,j}\\
v_{x,r} &= v_{x,i} - v_{x,j}, \qquad v_{y,r} = v_{y,i} - v_{y,j}
\end{aligned}
\end{equation}

We treat $Q_i$ as Player 1, the evader who wishes to avoid collision, and we treat $Q_j$ as Player 2, the pursuer, or disturbance, that wishes to cause a collision. In terms of the relative variables given in \eqref{eq:rel_var}, we have

\begin{equation}
\begin{aligned}
\dot{p}_{x,r}& = v_{x,r}, &\dot{p}_{y,r} &= v_{y,r} \\
\dot{v}_{x,r}& = u_{x,i} - u_{x,j}, &\dot{v}_{y,r} &= u_{y,i} - u_{y,j}\\
\end{aligned}
\end{equation}

We also augment \eqref{eq:rel_var} with the velocity of $Q_i$, given in \eqref{eq:rel_dyn_aug}, to impose a velocity limit on the quadrotor.

\begin{equation} \label{eq:rel_dyn_aug}
\begin{aligned}
\dot{p}_{x,r} &= v_{x,r}, &\dot{p}_{y,r} &= v_{y,r} \\
\dot{v}_{x,r} &= u_{x,i} - u_{x,j}, &\dot{v}_{y,r}&= u_{y,i} - u_{y,j}\\
\dot{v}_{x,i} &= u_{x,i}, &\dot{v}_{y,i} &= u_{y,i} \\
\end{aligned}
\end{equation}

\section{Liveness Controllers \label{sec:liveness}}
\subsection{Merging onto a Highway \label{subsec:highway_merge}}
We model the merging of a vehicle onto an air highway as a path planning problem, where we specify a target position and velocity whose magnitude (the speed) is given by the highway specification, and whose direction is along the highway. Thus, for a quadrotor, the objective would be to drive the system in \eqref{eq:dyn} to a specific state $\bar{x}_H=(\bar{p}_x, \bar{v}_x, \bar{p}_y, \bar{v}_y)$, or a small range of states defined by the set

@@

In this reachability problem, $\mathcal{L}_H$ is the target set, represented by the zero sublevel set of the function $l_H(x)$, which specifies the terminal condition of the HJB PDE that we need to solve. The solution we obtain, $V_H(t,x)$, is the implicit surface function representing the reachable set $\mathcal V_H(t)$; $V_H(-T,x)\le 0$, then, specifies the reachable set $\mathcal{V}_H(T)$, the set of states from which the system can be driven to the target $\mathcal{L}_H$ within a duration of $T$. This gives the algorithm for merging onto the highway:

\begin{enumerate}
\item Move towards $\bar{x}_H$ in a straight line, with some velocity, until $V_H(-T,x)\le 0$.
\item Apply the optimal control extracted from $V_H(-T,x)$ according to \eqref{eq:HJB_ctrl_syn} until $\mathcal{L}_H$ is reached.
\end{enumerate}

\subsection{Merging into a Platoon \label{subsec:platoon_merge}}
We again pose the merging of a quadrotor into a platoon on an air highway as a reachability problem. Here, we would like quadrotor $Q_i$ to merge onto the highway and follow another quadrotor $Q_j$ in a platoon. Thus, we would like to drive the system given by \eqref{eq:rel_dyn_aug} to a specific $\bar{x}_P = (\bar{p}_{x,r}, \bar{v}_{x,r}, \bar{p}_{y,r}, \bar{v}_{y,r})$, or a small range of relative states defined by the set

@@

The target set $\mathcal{L}_P$ is represented by the implicit surface function $l_P(x)$, which specifies the terminal condition of the HJI PDE \eqref{eq:HJIPDE}. The zero sublevel set of the solution to \eqref{eq:HJIPDE}, $V_P(-T,x)$, gives us the set of relative states from which $Q_i$ can reach the target and join the platoon following $Q_j$ within a duration of $T$. We assume that $Q_j$ moves along the highway at constant speed, so that $u_j(t)$ = 0. The following is a suitable algorithm for a quadrotor merging onto a highway and joining a platoon to follow $Q_j$:

\begin{enumerate}
\item Move towards $\bar{x}_P$ in a straight line, with some velocity, until $V_P(-T,x)\le 0$.
\item Apply the optimal control extracted from $V_P(-T,x)$ according to \eqref{eq:HJI_ctrl_syn} until $\mathcal{L}_P$ is reached.
\end{enumerate}

\subsection{Other Quadrotor Maneuvers}
Reachability was used in Sections \ref{subsec:highway_merge} and \ref{subsec:platoon_merge} for the relatively complex maneuvers. For the simpler maneuvers of traveling along a highway and following a platoon, we resort to simpler controllers described below.

\subsubsection{Traveling along a highway} \label{sec:travel_hwy}
We use a model-predictive controller (MPC) for traveling along a highway; this controller allows the leader to travel along a highway at a pre-specified speed. Here, the goal is for a leader quadrotor to track a constant-altitude path, defined as a curve $\bar{p}(s)$ parametrized by $s\in[0,1]$ in $p=(p_x, p_y)$ space (position space), while maintaining a velocity $\bar{v}(s)$ that corresponds to constant speed in the direction of the highway. Assuming that the initial position on the highway, $s_0=s(t_0)$ is specified, such a controller can be obtained from the following optimization problem over the time horizon $[t_0, t_1]$:

\begin{equation}
\begin{aligned}
\text{minimize } & \int_{t_0}^{t_1} \big\{\| p(t)-\bar{p}(s(t)) \|_2 + \\ 
&\qquad \| v(t) - \bar{v}(s(t)) \|_2 + 1-s \big\} dt \\
\text{subject to } & \dot{x} = f(x,u) \text{ where } f \text{ is given in \eqref{eq:dyn}} \\
& |u_x|, |u_y| \le u_\text{max}, |v_x|, |v_y| \le v_\text{max} \\
& s(t_0) = s_0, \dot{s} \ge 0
\end{aligned}
\end{equation}

If we discretize time, and assume that $\bar{p}(\cdot)$ is linear, then the above optimization is convex, and can be quickly solved.

\subsubsection{Following a Platoon} \label{sec:follow_platoon}
Follower vehicles use a feedback control law tracking a nominal position and velocity in the platoon, with an additional feed-forward term given by the leader's acceleration input; here, for simplicity, we assume perfect communication between the leader and the follower vehicles. This following law enables smooth vehicle trajectories in the relative platoon frame, while allowing the platoon as a whole to perform agile maneuvers by transmitting the leader's acceleration command $u_{P_1}(t)$ to all vehicles.

The $i$-th member of the platoon, $Q_{P_i}$, is expected to track a relative position in the platoon $r^i = (r_x^i,r_y^i)$ with respect to the leader's position $p_{P_1}$, and the leader's velocity $v_{P_1}$ at all times. The resulting control law has the form:
\begin{equation}\label{eq:follow}
u^i(t) = k_p \big[p_{P_1}(t) + r^i(t) - p^i(t) \big] + k_v\big[v_{P_1}(t) - v^i(t)\big] + u_{P_1}(t)
\end{equation}
for some $k_p,k_v>0$. The leader can modify the nominal position of vehicles in the platoon, for example to command the formation to turn. In particular, a simple rule for determining $r^i(t)$ in a single-file platoon is given for $Q_{P_i}$ as:
\begin{equation}\label{eq:nominal_pos}
r^i(t) = - (i-1) b \frac{v_{P_1}}{\|v_{P_1}\|_2}
\end{equation}
where $b$ is the spacing between vehicles along the platoon. and $\frac{v_{P_1}}{\|v_{P_1}\|_2}$ is the platoon leader's direction of travel.

\section{Safety Controllers \label{sec:safety}}
\subsection{Wrapping Reachability Around Existing Controllers}
A quadrotor can use a liveness controller when it is not in any danger of collision with other quadrotors or obstacles. If the quadrotor could potentially be involved in a collision within the next short period of time, it must switch to a safety controller. In this section, we will demonstrate how HJ reachability can be used to both detect imminent danger and synthesize a controller that guarantees safety within a specified time horizon. For our safety analysis, we will use the model in Equation \eqref{eq:rel_dyn_aug}.

We begin by defining the target set $\mathcal{L}_S$, which characterizes configurations in relative coordinates for which quadrotors $Q_i,Q_j$ are considered to be in collision:

@@

\subsection{Platoon Safety Guarantees \label{subsec:safety_guarantees}}
Under normal operations in a single platoon, each follower quadrotor $Q_i,i=P_2,\ldots,P_{N-1}$ in a platoon checks whether it is in the safety reachable set with respect to $Q_{P_{i-1}}$ and $Q_{P_{i+1}}$. So $\mathcal{Q}(i) = \{P_{i+1}, P_{i-1}\}$ for $i=P_2,\ldots,P_{N-1}$. Assuming there are no nearby quadrotors outside of the platoon, the platoon leader $Q_{P_1}$ checks safety against $Q_{P_2}$, and the platoon trailer $Q_{P_N}$ checks safety against $Q_{P_{N-1}}$. So $\mathcal{Q}(P_1)=\{P_2\}, \mathcal{Q}(P_N)=\{P_{N-1}\}$. When all quadrotors are using liveness controllers to perform their allowed maneuvers, no pair of quadrotors should be in an unsafe configuration if the liveness controllers are well-designed. However, occasionally a quadrotor $Q_k$ may behave unexpectedly due to faults or malfunctions, in which case it may come into an unsafe configuration with another quadrotor.

With our choice of $\mathcal{Q}(i)$ and the assumption that the platoon is in a single-file formation, some quadrotor $Q_i$ would get into an unsafe configuration with $Q_k$, where $Q_k$ is likely to be the quadrotor in front or behind of $Q_i$. In this case, a ``safety breach" occurs. Our synthesis of the safety controller guarantees that between every pair of quadrotors $Q_i,Q_k$, as long as $V_S(-t,x_i- x_k)>0$, $\exists u_i$ to keep $Q_i$ from colliding with $Q_k$ for a desired time horizon $t$, despite the worst case (an adversarial) control from $Q_k$. Therefore, as long as the number of ``safety breaches" is at most one, $Q_i$ can simply use the optimal control to avoid $Q_k$ and avoid collision for the time horizon of $t$. Since quadrotors in platoons are able to exit the current altitude range within a duration of $t_\text{internal}$, if we choose $t=t_\text{internal}$, the safety breach would always end before any collision can occur.

### 2016-03-03 docs/CDC 2015 platooning/final version/conclusion.tex
no message

% !TEX root = platooning.tex
\section{Conclusions and Future Work}
We considered single-file platoons of UAVs modeled by hybrid systems traveling along air highways. Using HJ reachability, we proposed liveness controllers for merging onto highways and merging into existing platoons, and wrapped a safety controller around liveness controllers to ensure no collision between the UAVs can occur. Under the assumption that faulty vehicles can descend after a pre-specified duration, our safety controller guarantees that no collisions will occur in a single altitude level as long as at most one safety breach occurs for each vehicle in the platoon. Additional safety breaches can be handled by multiple altitude ranges in the airspace. Our simulations show that by putting vehicles into single-file platoons, the likelihood of having multiple safety breaches is low, and conflicts involving a single malfunctioning UAV or intruder can be resolved in a single altitude level.

Immediate future work includes exploring different vehicle models, investigating algorithms for resolving multiple safety breaches within the same altitude, and more broadly related problems such as airspace structure and air highway placement.

>>>>

% !TEX root = platooning.tex
\section{Conclusions and Future Work}
We considered single-file platoons of UAVs modeled by hybrid systems traveling along air highways. Using HJ reachability, we proposed liveness controllers and built a safety controller around them to ensure no collision can occur from a single safety breach. Additional safety breaches can be handled by multiple altitude ranges in the airspace. Our simulations show that by putting vehicles into single-file platoons, the likelihood of having multiple safety breaches is low, and conflicts involving a single malfunctioning UAV or intruder can be resolved in a single altitude level.

### 2016-03-03 docs/CDC 2015 platooning/final version/results.tex
no message

Figure \ref{fig:normal2} illustrates the use of liveness and safety reachable sets using just two quadrotors to reduce visual clutter. The first quadrotor $Q_1$ (red disk) first travels in a straight line towards the highway merging point $\bar{x}$ (red circle) at $t=1.5$, because it is not yet in the liveness reachable set for merging onto the highway (red dotted boundary). When it is within the liveness reachable set boundary at $t=2.8$, it is ``locked-in" to the target state $\bar{x}_H$, and follows the optimal control in \eqref{eq:HJB_ctrl_syn} to $\bar{x}_H$. During the entire time, the $Q_1$ checks whether it may collide with $Q_2$ within a time horizon of $t_\text{external}$; we chose $t_\text{external}=3>t_\text{internal}$.

@@

Initially, $Q_{P_2}$ ($P_2=2$) is a follower and is outside all 3 safety reachable sets. At time $t=0.6$, $Q_2$ comes to the boundary of the safety set with respect to the intruder and therefore must apply the safety control law to avoid potential future collision. Thus it splits the original platoon and becomes the leader of a new platoon consisting of itself, $Q_3$ and $Q_4$. $Q_2$ keeps using the safety controller until it is safe with respect to the intruder again at $t=3$. After $t=3$, $Q_2$ is safe to use the liveness controller again to merge back onto the highway and join the original platoon as a follower. Note that during the entire time, $Q_2$ maintains safety against the intruder, $Q_1$ and $Q_3$ by always staying outside of all three safety reachable sets.

>>>>

Figure \ref{fig:normal2} illustrates the use of liveness and safety reachable sets using just two quadrotors to reduce visual clutter. The first quadrotor $Q_1$ (red disk) first travels in a straight line towards the highway merging point $\bar{x}$ (red circle) at $t=1.5$, because it is not yet in the liveness reachable set for merging onto the highway (red dotted boundary). When it is within the liveness reachable set boundary at $t=2.8$, it is ``locked-in" to the target state $\bar{x}_H$, and follows the optimal control in \eqref{eq:HJB_ctrl_syn} to $\bar{x}_H$. During the entire time, $Q_1$ checks whether it may collide with $Q_2$ within a time horizon of $t_\text{external}$; we chose $t_\text{external}=3>t_\text{internal}$.

@@

Initially, $Q_{P_2}$ ($P_2=2$) is a follower outside all 3 safety reachable sets. At $t=0.6$, $Q_2$ comes to the boundary of the safety set with respect to the intruder and must apply the safety control law to avoid potential future collision. Thus it splits from the original platoon and becomes the leader of a new platoon consisting of itself, $Q_3$ and $Q_4$. $Q_2$ keeps using the safety controller until it is safe with respect to the intruder again at $t=3$. After $t=3$, $Q_2$ is safe to use the liveness controller again to merge back onto the highway and join the original platoon. Note that during the entire time, $Q_2$ maintains safety against the intruder, $Q_1$ and $Q_3$ by always staying outside of all three safety reachable sets.

### 2016-03-03 docs/CDC 2015 platooning/initial version/introduction.tex
no message

% !TEX root = main.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications. Through projects such as Amazon Prime Air \cite{PrimeAir} and Google Project Wing \cite{ProjectWing}, companies are looking to send UAVs into the airspace to not only deliver commerical packages, but also for important tasks such as aerial surveillance, emergency supply delivery, and search and rescue. In the future, the applications of UAVs are only limited by human imagination.

As a rough estimate, suppose in a city of 2 million people, each person requests a drone delivery every 2 months on average and each delivery requires a 30 minute trip for the drone. This would equate to thousands of vehicles simultaneously in the air just from package delivery services. Applications of UAVs extend beyond package delivery; they can also be used, for example, to provide supplies or to respond to disasters in areas that are difficult to reach but require prompt response \cite{Debusk10,Tornado16}. As a result, government agencies such as the Federal Aviation Administration (FAA) and National Aeronautics and Administration (NASA) are also investigating air traffic control for autonomous vehicles in order to prevent collisions among potentially numerous UAVs \cite{FAA13, NASA16}.

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under bounded disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving a small number of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

In order to accommodate potentially thousands of vehicles simultaneously flying in the air, additional structure is needed to allow for tractable analysis and intuitive monitoring. An air highway system on which platoons of vehicles travel accomplishes both goals. However, many details of such a concept need to be addressed. Due to the flexibility of placing air highways compared to building ground highways in terms of highway location, even the problem of air highway placement can be a daunting task. To address this, in the first part of this paper, we propose a flexible and computationally efficient way based on \cite{Sethian96} to perform optimal air highway replacement given an arbitrary cost map that captures the desirability of having UAVs fly over any geographical location. We demonstrate our method using the San Francisco Bay Area as an example. Once air highways are in place, platoons of UAVs can then fly in fixed formations along the highway to get from origin to destination.

@@

Previous analysis of a large number of vehicles typically do not provide liveness and safety guarantees to the extent that HJ reachability does; however, HJ reachability typically cannot be used to tractably analyze a large number of vehicles. In the second part of this paper, we propose organizing UAVs into platoons, which provide structure that allows pairwise safety guarantees from HJ reachability to translate to safety guarantees for the whole platoon. With respect to platooning, we first propose a hybrid systems model of UAVs in platoons to establish the modes of operation needed for our platooning concept. Then, we show how reachability-based controllers can be synthesized to enable vehicles to successfully perform mode switching, as well as prevent dangerous configurations such as collisions. Finally, we show several simulations to illustrate the behavior of vehicles in various scenarios and demonstrate the guarantees provided by HJ reachability.

>>>>

% !TEX root = platooning.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications. Through projects such as Amazon Prime Air and Google Project Wing, companies are looking to send UAVs into the airspace to deliver packages \cite{PrimeAir,ProjectWing}. As a rough estimate, suppose in a city of 2 million people, each person requests a drone delivery every 2 months on average and each delivery requires a 30 minute trip for the drone. This would equate to thousands of vehicles simultaneously in the air. As a result, government agencies such as the Federal Aviation Administration (FAA) and National Aeronautics and Administration (NASA) of the United States are also investigating air traffic control for autonomous vehicles in order to prevent collisions among potentially numerous UAVs \cite{FAA13}. Applications of UAVs extend beyond package delivery; they can also be used, for example, to provide supplies or to firefight in areas that are difficult to reach but require prompt response \cite{Debusk10}.

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving a small number of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

@@

Previous analysis of a large number of vehicles typically do not provide liveness and safety guarantees to the extent that HJ reachability does; however, HJ reachability typically cannot be used to tractably analyze a large number of vehicles. %In the context of HJ reachability, putting vehicles into platoons is desirable because of the additional structure that platoons impose on its members. With additional structure, pairwise safety guarantees of vehicles can be more easily translated into safety guarantees of all the vehicles in the platoon. 
In this paper, we attempt to reconciliate this trade-off by assuming a single-file platoon, which provides structure that allows pairwise safety guarantees from HJ reachability to translate to safety guarantees for the whole platoon. We first propose a hybrid systems model of UAVs in platoons to capture this structure. Then, we show how HJ reachability can be used to synthesize \textit{liveness controllers} that enable vehicles to reach a set of desired states, and wrap \textit{safety controllers} around the liveness controllers in order to prevent dangerous configurations such as collisions. Finally, we show simulation results of quadrotors forming a platoon, a platoon responding to a malfunctioning member, and a platoon responding to an outside intruder to illustrate the behavior of vehicles in these scenarios and demonstrate the guarantees provided by HJ reachability.

%Motivation:
%\begin{itemize}
%\item Applications of UAVs, potential numbeurs
%\item Importance of safety guarantees
%\item Computation complexity
%\end{itemize}
%
%Related work:
%\begin{itemize}
%\item Platooning references: lack of (?) safety guarantees
%\begin{itemize}\item\textcolor{blue}{Mention PATH papers \& Lygeros thesis}\end{itemize}
%\item HJI, safety guarantees, limited by dimensionality
%\end{itemize}
%
%Summary of results
%\begin{itemize}
%\item Hybrid systems model of vehicles
%\item Reachability guarantees wrapped around existing methods
%\item Reachability offers flexibility in terms of design
%\item Illustrative platoon functions
%\end{itemize}
